TODO: title

### 2014-05-04 paper/main.tex
outlined or written first two sections, secction three is written, but needs extensive review, possible re-outlining.

  % bridge from past work to our work.
    % point out holes (evaluating standard algorithms for little to no ratings, wholeistic evaluation)
    % point out that other approaches than smart object ordering exist, and we need to be able to compare.
    % cite the comparison work at CHI.
    % we propose to evaluate 
  While past work has covered the new user rating survey as a solution to new user cold start, much less work has been done comparing other strategies.

@@

  \subsection*{Metrics}
  We will discuss a total of eleven metrics.
%%%%%%%%%%%%%%%%%%%%%%
  Some of these metrics will prove to be redundant, and future work should be able to use a much smaller list.
  The metrics are listed in table \ref{tbl:metrics}.
  \begin{table}[ht!]
    \centering
    \begin{tabular}{|p{7em}|p{16em}|}
      \hline
      \multicolumn{2}{|c|}{{\bf Accuracy Metrics}} \\\hline
      RMSE                   & Standard prediction accuracy metric. \\\hline
      \hline
      \multicolumn{2}{|c|}{{\bf Recommendation Metrics}} \\\hline
      nDCG                   & Standard ranking quality evaluation.\\\hline
      Precision@20           & Count of recommended items in the test set and rated 4.0 or higher.\\\hline
      MAP@20                 & Average of precision at each rank one through twenty.\\\hline
      Fallout@20             & Count of recommended items in the test set and rated 2.0 or lower.\\\hline
      SeenItems@20           & Count of recommended items in the test set.\\\hline
      MeanRating@20          & Average rating of recommended items that are in the test set.\\\hline
      RMSE@20                & RMSE computed only over those items that were recommended.\\\hline
      \hline
      \multicolumn{2}{|c|}{{\bf Other Metrics}} \\\hline
      Average\-Popularity@20 & The average number of users in the training set that have seen recommended items.\\\hline
      AILS@20                & The average pairwise similarity between recommended items. \\\hline
      Spread@20              & The entropy of the distribution of recommended items for users in the test set.\\\hline
    \end{tabular}
    \caption{TODO:}
    \label{tbl:metrics}
  \end{table}

  All metrics evaluating recommendations (all but RMSE and nDCG) are based on the top 20 predicted items for the user, excluding items in the training set.
  The AILS@20 metric is based on the ILS metric introduced by ziegler et al \cite{ziegler}.
  AILS@N is simply a rescaling of ILS to be scale free with regards to N, this was done to make the metric easier to interpret, by putting it on the same scale as similarity values.
  The choice of 20 items for the evaluations was arbitrary, testing with other values showed similar results, except on one metric.
  This issue will be addressed later in the results section.

%This is simply a rescaling of the ILS metric from \cite{ziegler}. Similarity is computed by subtracting the average item rating from the item vectors for each item and then computing the cosine similarity.\\\hline

  % motivate many metrics.
  % metrics (expected / good behavior for each metric (I.E. this should go down, any reasonable algorithm should always beet baseline)) 
    % accuracy metrics 
      % RMSE
    % recommendation performance metrics (list subject to change)
      % ndcg
      % MAP@N (not yet measured)
      % precision@N
      % fallout @ N
      % RMSE of recommended items
      % (note that we looked at MRR and recall, but found them to behave the same).
    % other properties
      % recommendation average popularity
      % recommendation diversity
      % spread (entropy of recommendations) (related to recommendation coverage)

  \subsection*{Dataset}
  % dataset
    % # user
    % # items
    % distributions of # ratings per user (important: min 20)

  \subsection*{Evaluation Procedure}

  % subsampling strategy
    % describe normal crossfold solution
    % describe issues with solution
    % describe our crossfold then subsample solution.

>>>>

  % bridge from past work to our work.
    % point out holes (evaluating standard algorithms for little to no ratings, wholeistic evaluation)
    % point out that other approaches than smart object ordering exist, and we need to be able to compare.
    % cite the comparison work at CHI.
  While past work has covered the new user rating survey as a solution to new user cold start, much less work has been done comparing other strategies.

  % Temporal evaluation
    % similar in-so-much as it can provide similar answers (growth over time), 
    % different in-so-much as it tends to be used for just system growth, but it could be used for user growth.
  Some researchers have taken to perfforming temporal evaluations of recommender systems, looking at accuracy metrics as a function of the lifetime of a system.

  % inspired by this work...

@@

  \subsection*{Metrics}
  We will discuss a total of eleven metrics.
%%%%%%%%%%%%%%%%%%%%%%
  Some of these metrics will prove to be redundant, and future work should be able to use a much smaller list.
  The metrics are listed in table \ref{tbl:metrics}.
  \begin{table}[ht!]
    \centering
    \begin{tabular}{|p{7em}|p{16em}|}
      \hline
      \multicolumn{2}{|c|}{{\bf Accuracy Metrics}} \\\hline
      RMSE                   & Standard prediction accuracy metric. \\\hline
      \hline
      \multicolumn{2}{|c|}{{\bf Recommendation Metrics}} \\\hline
      nDCG                   & Standard ranking quality evaluation.\\\hline
      Precision@20           & Count of recommended items in the test set and rated 4.0 or higher.\\\hline
      MAP@20                 & Average of precision at each rank one through twenty.\\\hline
      Fallout@20             & Count of recommended items in the test set and rated 2.0 or lower.\\\hline
      SeenItems@20           & Count of recommended items in the test set.\\\hline
      MeanRating@20          & Average rating of recommended items that are in the test set.\\\hline
      RMSE@20                & RMSE computed only over those items that were recommended.\\\hline
      \hline
      \multicolumn{2}{|c|}{{\bf Other Metrics}} \\\hline
      Average\-Popularity@20 & The average number of users in the training set that have seen recommended items.\\\hline
      AILS@20                & The average pairwise similarity between recommended items. \\\hline
      Spread@20              & The Shannon's entropy of the distribution of recommended items for users in the test set.\\\hline
    \end{tabular}
    \caption{TODO:}
    \label{tbl:metrics}
  \end{table}

  All metrics evaluating recommendations (all but RMSE and nDCG) are based on the top 20 predicted items for the user, excluding items in the training set.
  The choice of 20 items for the evaluations was arbitrary, testing with other values showed similar results, except on one metric.
  This issue will be addressed later in the results section.

  % Provide computation details for AILS@N
  Most of our metrics are standard techniques, or easily commutable, therefore we will only comment on the AILS, and spread metrics.
%%%%%%%%%%%%%%%%%%%%%%%%%
  The AILS@20 metric is based on the ILS metric introduced by ziegler et al \cite{ziegler}.
  AILS@N is simply a rescaling of ILS to be scale free with regards to N, this was done to make the metric easier to interpret, by putting it on the same scale as similarity values.

  The spread metric is computed by computing recommendations for each test user.
  Let $C(I)$ be the count of how many times item $I$ showed up in the recommendations.
  Let $P(I) = C(I) / \sum_I C(I)$ be the probability that a random recommendation is item $I$.
  Finally we define spread as the Shannon's entropy of $P(I)$, $Spread = -\sum_I P(I) \log(P(I))$.
%%%%%%%%%%%%%%%%%%%%%%%%%%

  \subsection*{Dataset}
  % dataset
    % # user
    % # items
    % distributions of # ratings per user (important: min 20)
    % establish range of evaluation.
  For this evaluation we will use the MovieLens 1M dataset.

  % movielens system
      % silly cold start
      % randomization
  The MovieLens system uses a new user rating survey to allow users to enter the system.

  \subsection*{Evaluation Procedure}
  % subsampling strategy
  % describe normal crossfold solution
  The natural way to perform this evaluation is with a modified user crossfolding strategy.

  % describe issues with solution
  Unfortunately, we found that this strategy can causes biases in our evaluation.

  % describe our crossfold then subsample solution.
  Our solution to this bais is to downsample an existing crossfold.
